Reproduction frequency and offspring survival decline in elephant seals past prime age

### 2024-08-13 paper/paperv3.qmd

@@

$$
\pi_{1a} = \frac{\sum_{a=4}^{20} b(a)n(a)}{\sum_{a=4}^{20} n(a)}
$$

@@

$$
\pi_{2a} = \frac{\sum_{a=4}^{11} b(a) n(a) + \sum_{a=12}^{20} b(11) n(a)}{\sum_{a=4}^{20}n(a)}
$$

@@

$$
\pi_{1b} = \frac{\sum_{a=4}^{20} b(a) n(a) s(a)}{\sum_{a=4}^{20} n(a)}
$$

$$
\pi_{2b} = \frac{\sum_{a=4}^{11} b(a) n(a) s(a) + \sum_{a=12}^{20} b(11) n(a) (s11)}{\sum_{a=4}^{20}n(a)}
$$

@@

str_glue("
delta pi a {round(100 * (pi1a - pi2a) / pi2a, 2)}%,
delta pi b {round(100 * (pi1b - pi2b) / pi2b, 2)}%")

>>>>

## Nursing duration

What's the mean and 95% CI duration of nursing? [@reiter1981] provide the following information in Table III:

+-------+----------+----------+----------+----------+----------+----------+----------+
| 2     | 3        | 4        | 5        | 6        | 7        | 8        | 9        |
+=======+==========+==========+==========+==========+==========+==========+==========+
| 22±0  | 23.7±1.6 | 26.6±1.6 | 26.1±1.9 | 26.5±2.2 | 27.1±1.2 | 28.2±0.8 | 28.5±0.7 |
|       |          |          |          |          |          |          |          |
| \(1\) | \(9\)    | \(24\)   | \(12\)   | \(6\)    | \(8\)    | \(5\)    | \(2\)    |
+-------+----------+----------+----------+----------+----------+----------+----------+

Where the column names are the age of the seal in years, and row values indicate mean±SD (n). To get the mean and 95% CI duration of nursing, we pooled the samples for mothers age 4 and up (i.e., excluded 2- and 3-year olds).

The formulas we used for pooling the variances come from [@oneill2014].

```{r}
#| label: pooled-nursing

# N, mean, and SD from Reiter et al. (1981) Table III for 4+ yo mothers
n <- c(24, 12, 6, 8, 5, 2)
m <- c(26.6, 26.1, 26.5, 27.1, 28.2, 28.5)
s <- c(1.6, 1.9, 2.2, 1.2, 0.8, 0.7)

# Function for pooling two samples
pool_samples <- function(sample1, sample2) {
  # sample1 and sample2 must be numeric vectors of length 3 with names v, m, and n
  is_valid <- function(x) {
    is.numeric(x) &&
      is.vector(x) &&
      (length(x) == 3) &&
      all(c("v", "m", "n") %in% names(x))
  }

  if (!(is_valid(sample1) && is_valid(sample2))) 
    stop("sample1 and sample2 must be numeric vectors of length 3 with names v, m, and n (variance, mean, and number of elements)")

  v1 <- sample1["v"]
  m1 <- sample1["m"]
  n1 <- sample1["n"]
  v2 <- sample2["v"]
  m2 <- sample2["m"]
  n2 <- sample2["n"]

  n3 <- n1 + n2
  m3 <- weighted.mean(c(m1, m2), c(n1, n2))
  v3 <- 1 / (n1 + n2 - 1) * ((n1 - 1) * v1 + (n2 - 1) * v2 + n1 * n2 / (n1 + n2) * (m1 - m2)^2)
  result <- c(v3, m3, n3)
  names(result) <- c("v", "m", "n")
  result
}

# Pool the samples iteratively using reduction
samples <- pmap(list(n, m, s), \(.n, .m, .s) c(v = .s^2, m = .m, n = .n))
pooled_samples <- reduce(samples, pool_samples)

sprintf("%0.1f (95%%CI: %0.1f - %0.1f)", 
        pooled_samples["m"],
        pooled_samples["m"] - 1.96 * sqrt(pooled_samples["v"]),
        pooled_samples["m"] + 1.96 * sqrt(pooled_samples["v"]))
```

@@

$$
\pi_{1a} = \frac{\sum_{a=4}^{20} b_a n_a}{\sum_{a=4}^{20} n_a}
$$

@@

$$
\pi_{2a} = \frac{\sum_{a=4}^{11} b_a n_a + \sum_{a=12}^{20} b_{11} n_a}{\sum_{a=4}^{20}n_a}
$$

@@

$$
\pi_{1b} = \frac{\sum_{a=4}^{20} b_a n_a s_a}{\sum_{a=4}^{20} n_a}
$$

$$
\pi_{2b} = \frac{\sum_{a=4}^{11} b_a n_a s_a + \sum_{a=12}^{20} b_{11} n_a s_{11}}{\sum_{a=4}^{20} n_a}
$$

@@

str_glue("
delta pi a {round(100 * (pi1a - pi2a) / pi2a, 2)}%,
delta pi b {round(100 * (pi1b - pi2b) / pi2b, 2)}%")
```
